Pyrovelocity PyPI user notebook template

### 2024-03-21 scripts/examples/pypi_template/pyrovelocity_pypi_user_template.qmd
fix(pypi_template): add blank header cells Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>
fix(pypi_template): sync with library updates Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

@@

@@

Installation should take less than **5 minutes**. 
It involves checking if the notebook is running in colab, in which case it is certain that you will need to install or reinstall pyrovelocity.
Otherwise, it is assumed that the user will have installed pyrovelocity in the same environment in which the jupyter kernel is running.

@@

@@

if IN_COLAB:
    colab_release_tag = os.getenv("COLAB_RELEASE_TAG", None)
    print(f"Google Colab release: {colab_release_tag}")
    setup_pyrovelocity()
else:
    print("This notebook is probably not running in Google Colab")
```

@@

This is the same for the pyro package in case there was an issue with pyrovelocity install that did not affect another package.

@@

```{python}
#| label: print-pyrovelocity-docstring-1
pyrovelocity.help.internal_help(pyrovelocity) # ?pyrovelocity # to open in side panel tab for reference
```

@@

```{python}
#| label: import-notebook-dependencies
from importlib import reload

import yaml
from IPython.display import Image
# from IPython.display import display

import pyrovelocity.utils
import pyrovelocity.workflows.main_workflow
reload(pyrovelocity.utils)
reload(pyrovelocity.workflows.main_workflow)

from pyrovelocity.utils import print_config_tree
from pyrovelocity.utils import print_docstring
from pyrovelocity.workflows.main_workflow import download_data
from pyrovelocity.workflows.main_workflow import postprocess_data
from pyrovelocity.workflows.main_workflow import preprocess_data
from pyrovelocity.workflows.main_workflow import summarize_data
from pyrovelocity.workflows.main_workflow import train_model
```

@@

::: {.callout-note collapse=true title="Local caching"}
During local execution, all of the functions imported from the `pyrovelocity.workflows.main_workflow` module will cache their results. This means that if you run the same task function with the same arguments, it will not recompute the result, but just establish the presence of a cache hit. This is useful for avoiding redundant computation, but if you want to re-execute the function for testing, experimentation or other purposes. To avoid this, you can use the `pyrovelocity.utils.clear_local_cache` function to clear the cache. If you add the following block of code, the cache will be cleared whenever it is executed:

@@

@@

```{python}
#| label: execute-summarize-data-task
#| output: false
dataset_summary = summarize_data(
  preprocess_data_args=pancreas_configuration.preprocess_data,
  postprocessing_outputs=postprocessing_outputs,
  training_outputs=model_output,
)
```

@@

### Data set review 
```{python}
#| label: import-ipython-display
#| echo: true
from IPython.display import Image, display
```

@@

display(Image(filename=f"{dataset_summary.data_model_reports.path}/fig2_part1_plot.pdf.png"))
```

@@

display(Image(filename=f"{dataset_summary.data_model_reports.path}/rainbow.pdf.png"))
```

@@

display(Image(filename=f"{dataset_summary.data_model_reports.path}/fig2_part2_plot.pdf.png"))
```

@@

```{python}
#| label: fig-manually-generated-rainbowplot
#| echo: true
#| fig-cap: "Manually generated rainbow plot." 
with SuppressOutput():
  pv.plots.rainbowplot(
      volcano_data=volcano_data,
      adata=adata,
      posterior_samples=posterior_samples,
      genes=putative_marker_genes,
      data=["st", "ut"],
      basis="umap",
      cell_state="clusters",
  )
```

@@

```{python}
#| label: print-pyrovelocity-docstring-2
pv.help.internal_help(pv)
```

>>>>

```{python}
#| label: blank-cell-01
#| code-fold: true
#| output: false
print("This cell is intentionally left blank.") 
```

@@

```{python}
#| label: blank-cell-02
#| code-fold: true
#| output: false
print("This cell is intentionally left blank.") 
```

@@

Installation should take less than **10 minutes**. If you are running in [Google Colab][colab-more-memory], then even when running with the `PYROVELOCITY_TESTING_FLAG` set to `True`,
if you have [access to *High-RAM* or *T4 GPU* runtimes][colab-more-memory] this can significantly speed up even this testing of the notebook.
For a complete analysis with `PYROVELOCITY_TESTING_FLAG` disabled, at least *High-RAM* or *T4 GPU* runtimes are essentially required to execute in a reasonable
amount of time.

This section involves checking if the notebook is running in colab, in which case it is certain that you will need to install or reinstall pyrovelocity.
Otherwise, it is assumed that the user will have installed pyrovelocity in the same environment in which the jupyter kernel is running.

@@

[colab-more-memory]: https://colab.research.google.com/notebooks/pro.ipynb#scrollTo=65MSuHKqNeBZ

@@

if IN_COLAB:
    colab_release_tag = os.getenv("COLAB_RELEASE_TAG", None)
    print(f"Google Colab release: {colab_release_tag}")
    setup_pyrovelocity()

    from google.colab import output
    output.enable_custom_widget_manager()
else:
    print("This notebook is probably not running in Google Colab")
```

@@

In case there is an issue with `pyrovelocity` installation, the above output should be analogous to that shown below for the pyro package.

@@

```{python}
#| label: print-pyrovelocity-docstring-1
pyrovelocity.utils.internal_help(pyrovelocity) # ?pyrovelocity # to open in side panel tab for reference
```

@@

```{python}
#| label: import-notebook-dependencies
import yaml

import pyrovelocity.utils
import pyrovelocity.workflows.main_workflow
from pyrovelocity.utils import print_config_tree
from pyrovelocity.utils import print_docstring
from pyrovelocity.workflows.main_workflow import download_data
from pyrovelocity.workflows.main_workflow import postprocess_data
from pyrovelocity.workflows.main_workflow import preprocess_data
from pyrovelocity.workflows.main_workflow import summarize_data
from pyrovelocity.workflows.main_workflow import train_model
```

::: {#nte-reload-modules .callout-note collapse=true title="Reload modules when installing in editable mode"}
If you install the package in editable mode with `pip install -e` for development purposes, you may find it helpful to use

```python
from importlib import reload

reload(pyrovelocity.utils)
reload(pyrovelocity.workflows.main_workflow)
```

We leave this out of this notebook since it is intended as a usage example.
:::

@@

::: {#nte-local-caching .callout-note collapse=true title="Local caching"}
During local execution, all of the functions imported from the `pyrovelocity.workflows.main_workflow` module will cache their results. This means that if you run the same task function with the same arguments, it will not recompute the result, but just establish the presence of a cache hit. This is useful for avoiding redundant computation, but if you want to re-execute the function for testing, experimentation or other purposes. To avoid this, you can use the `pyrovelocity.utils.clear_local_cache` function to clear the cache. If you add the following block of code, the cache will be cleared whenever it is executed:

@@

::: {#nte-suppressoutput .callout-note collapse=true title="Ignore or remove SuppressOutput"}
You can safely remove and dedent code executed in the context of `SuppressOutput` in a notebook execution environment.

The `SuppressOutput` context manager is used to suppress some of the output of the function calls placed in its context. This is purely used as a work-around for compatibility with multiple output formats such as html and pdf and it will be removed in the future. It is not necessary for the execution of the `postprocess_data` function in general or for execution in a notebook in particular.
:::

@@

```{python}
#| label: execute-summarize-data-task
#| output: false
from IPython.utils import io

with io.capture_output() as captured:
  dataset_summary = summarize_data(
    preprocess_data_args=pancreas_configuration.preprocess_data,
    postprocessing_outputs=postprocessing_outputs,
    training_outputs=model_output,
  )
```

::: {#nte-find-figures .callout-note collapse=true title="Path to figure files generated by summarize_data"}
The call to `summarize_data` generates a large number of plots which are saved to the `dataset_summary.data_model_reports.path`. This currently resolves to a subfolder of `./reports`. The name of the subfolder joins the data set name with the name of the model type. We capture the plot outputs to prevent them from generating a very long cell output. We review several of the plots in @sec-data-set-review. The remaining plots are available in the `dataset_summary.data_model_reports.path` folder. If you cannot locate it in the file browser, you can execute

```python
print(dataset_summary.data_model_reports.path)
```

after the cell containing the call to summarize_data completes to verify the path where the plots are located.
:::

@@

### Data set review {#sec-data-set-review}
```{python}
#| label: import-ipython-display
#| echo: true
from IPython.display import Image, display
```

@@

display(Image(filename=f"{dataset_summary.data_model_reports.path}/vector_field_summary_plot.pdf.png"))
```

@@

display(Image(filename=f"{dataset_summary.data_model_reports.path}/gene_selection_rainbow_plot.pdf.png"))
```

@@

display(Image(filename=f"{dataset_summary.data_model_reports.path}/gene_selection_summary_plot.pdf.png"))
```

@@

```{python}
#| label: fig-manually-generated-rainbowplot
#| echo: true
#| fig-cap: "Manually generated phase portrait, temporal dynamics, and shared time correlation rainbow plot." 
with SuppressOutput():
  pv.plots.rainbowplot(
      volcano_data=volcano_data,
      adata=adata,
      posterior_samples=posterior_samples,
      genes=putative_marker_genes,
      data=["st", "ut"],
      basis="umap",
      cell_state="clusters",
  )
```

See @nte-suppressoutput if you need to know what `SuppressOutput` is doing.

@@

```{python}
#| label: print-pyrovelocity-docstring-2
pv.utils.internal_help(pv)
```
